### 2015-06-17 text/ms.Rmd
Edit paper

What can we do when multiple ecological models give different impressions of population status? Any single model is unlikely to perform best in all scenarios. 
However, ensemble methods from the fields climate science and machine learning provide a powerful tool to leverage information across multiple ecological models. 
Here we evaluate the potential for ensemble methods to improve estimates of fish population status. 
We fit four data-limited models of population and exploitation dynamics and combine their estimates of the mean and slope of biomass at maximum sustainable yield (\bbmsy\\) with four ensemble models: a simple average, a linear model, and two machine-learning methods---random forests and boosted regression trees. 
We build our ensembles on a simulated dataset of 5760 fish stocks and test our models with cross-validation and a global database of `r ram_stocks_n` stock assessments. 
We find that ensemble models substantially improve estimates of population status and trajectory. 
Machine-learning ensembles performed the best: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, correlation improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. 
We found similar improvements when predicting the slope of status. 
When the simulation-trained ensembles were applied to global fish stocks, the best ensembles still improved all dimensions of performance over the next-best individual model. 
Ensemble models have potential to improve estimates of population status and trajectory across many taxa beyond fish. 
However, ensemble methods must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations they are applied to.

@@

There are often multiple ecological models we can apply to a system to assess
population status and trajectory. Models might incorporate different data
types, assume alternate population dynamics, or make contrasting assumptions
about the starting state of a system. These models may suggest different
conclusions about population status and trajectory and we often do not know
which model is best. How can we reconcile multiple models to make robust
management decisions about ecological resources?

Ensemble methods, which combine the output from multiple models as predictors
in a new 'ensemble' model, are one solution to this problem. A simple ensemble
model might take the average of multiple model outputs, perhaps weighted by
some goodness of fit. A complex ensemble model might also take into account the
interactions between model outputs and additional covariates. Ensemble models
are popular in other fields including public health [e.g. @thomson2006],
agriculture [e.g. @cantelaube2005], climate science [e.g. @murphy2004;
@tebaldi2007; @pierce2009], and machine learning [e.g. @dietterich2000]. We
think they have tremendous opportunity in ecological resource management.

In fisheries science, a common problem is estimating the status and trajectory
of an exploited fish population. For the majority of fish stocks, we have
limited biological information and stock status isn't known [@fao2014]. In
recent years, a number of methods have been proposed to derive population
status based on the limited information that is broadly available: fisheries
catch time series and basic knowledge about the productivity of the species.
Recently, @rosenberg2014 investigated the ability for these models to estimate
population status through a large-scale simulation model. A key finding was
that these models frequently disagree on population status and any one model
has relatively poor predictive ability on average (e.g. Fig.
\ref{motivate}).

@@

<!--TODO: github flr repo not available-->

We first developed and tested ensemble models with a fully factorial simulated
dataset [@rosenberg2014]. In summary, the simulation models included three life
histories: small pelagic, demersal, large pelagic; three levels of how depleted
biomass was at the start of the dataset compared to carrying capacity: 100%,
70%, and 40% of carrying capacity; and four harvest dynamics: a constant
harvest rate, a harvest rate that is coupled with biomass to mimic an
open-access single-species fishery, a scenario where harvest rate increases to
a fixed level, a 'roller-coaster' scenario where the harvest rate increases,
maintains a steady level and decreases to a lower steady level. Process noise
was introduced to the models through two levels of multiplicative recruitment
variability: $N(0, 0.2)$ and $N(0, 0.6)$, and that process noise was either
uncorrelated or had a first-order autoregressive correlation of 0.6. The
simulation included a scenario without observation error and with
multiplicative observation error around catch at $N(0, 0.2)$.

@@

We also tested our ensemble models on the RAM Legacy Stock Assessment Database
[@ricard2012]. Our analysis of the stock-assessment database was based on
version XX downloaded on XX. After removing stocks for which the models
described below did not converge, this database included `r ram_stocks_n` from
XX countries across XX taxonomic orders.

@@

We fit four individual data-limited models to estimate \bbmsy\\. Three of the
models are mechanistic and based generally on @schaefer1954 biomass
dynamics of the form:

@@

- *COM-SIR* (catch-only-model with sample-importance-resampling) is a coupled
  harvest-dynamics model developed by @vasconcellos2005. Biomass is assumed to
  follow a Schaefer model and harvest dynamics are assumed to follow a logistic
  model. The model is fit with a sample-importance-sampling algorithm developed
  in @rosenberg2014.

- *CMSY* (catch-MSY) implements a stock-reduction analysis with Schaefer
  biomass dynamics [@martell2013]. It requires a prior distribution for $r$ and
  assigns a prior to the relative proportion of biomass at the end compared to
  unfished biomass (depletion) based on the percentage of maximum catch at the
  end of the time series.

- *SSCOM* (state-space catch-only-model) TODO: Jim -- can you give a ~2 sentence summary?

- *mPRM* (modified panel regression model) is a modified version of the
  panel-regression model used in @costello2012. Unlike the other models
  considered, mPRM is empirical and not mechanistic --- it uses the RAM Legacy
  Stock Assessment database to fit a linear model to a series of
  characteristics of the catch time series and stock with stock-assessed
  \bbmsy\\ as the response. The model used in this paper and @rosenberg2014 is
  modified from the original in that it condenses the life-history categories
  into three categories to match the simulated dataset. 
  <!--TODO: what are we doing with the RAM stocks again?-->
  <!--TODO: bias adjustment?-->
  <!--TODO: build in fig:didactic -->

@@

An ensemble model can combine models through anything from a simple average to
a complex non-linear model. Here we choose four ensemble models to compare: an
average, a linear model with two-way interactions, a random forest, and boosted
regression tree. The general form of our models is as follows:

@@

\noindent where $L$ is a generalized loss function, $f$ is some generalized
regression function, $\hat{b}_\mathrm{CMSY}$ is the prediction from the CMSY
model etc. (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the
spectral density of the scaled catch series at frequencies of 0.2 and 0.05, $b$
is the true value the model is fitted to, and $\theta$ is the set of parameters
for the regression function. For all ensemble models of mean \bbmsy\\, which is
bounded at zero, we fit the models in log space and exponentiated the
predictions to derive a median estimate of \bbmsy\\. For the estimates of
\bbmsy\\ slope, we fit ensemble models on the natural untransformed scale.

@@

Applied to the simulated dataset of known status, the individual models had variable success at recovering the mean and slope of \bbmsy\\ in the last five years (hereafter used interchangeably with 'status' and 'trajectory', respectively). 
All models exhibited a high degree of scatter around the 1:1 line of perfect status prediction (Fig. \ref{hexagon}). 
CMSY exhibited bimodal predictions of status but had the best rank-order correlation and accuracy scores (Fig. \ref{performance}a).
COM-SIR and SSCOM both correctly identified a number of low status stocks, but frequently predicted a high status when status was in fact low (Fig. \ref{hexagon}b, c). 
mPRM had relatively poor ability to predict status for the simulated dataset (Fig. \ref{hexagon}d). 
There was generally little correlation between true and predicted trajectory for any of the individual models besides SSCOM (Figs \ref{scatter-sim-slope}a--d, \ref{performance-sim-slope}).

@@

The ensemble models were able to improve the predictions of status by exploiting the best properties of individual models, the covariance between individual models, and interactions with other covariates.
For example, SSCOM had strong predictive ability when it predicted low \bbmsy\\ (Fig. \ref{hexagon}c, Fig. \ref{partial-sim}c) and CMSY predictions were approximately linearly related to \bbmsy\\ within the low and high bimodal prediction clusters (Fig. \ref{partial-sim}).
Ensembles also exploited the covariance between individual model predictions.
For instance, both the linear model and GBM ensemble suggest that if mPRM and SSCOM predict high status, the true status also tends to be high (Figs \ref{lm-coefs}, \ref{partial-2d-sim}l).
The addition of spectral density covariates helped the ensemble models correctly predict higher status values (Fig. \ref{hexagon}g, h) by exploiting interactions between the spectral density predictors and the individual model predictions (Fig. \ref{partial-2d-sim}).
Although the additional covariates improved ensemble fit, performance of the ensembles was only marginally degraded by removing these covariates (Fig. \ref{hexagon} vs. Fig. \ref{hexagon-sim-basic}) TODO NEW FIG THAT QUANTIFIES THIS.

>>>>

What can we do when multiple ecological models suggest different conclusions about population status? Any single model is unlikely to perform best under all scenarios. Ensemble methods, commonly used in fields such as climate science, provide a powerful tool to leverage information across multiple ecological models. Here we evaluate the potential for ensemble methods to improve estimates of fish population status and trajectory. We fit four data-limited models of population and exploitation dynamics and combine their estimates of the mean and slope of biomass at maximum sustainable yield (\bbmsy\\) with four ensemble models: a simple average, a linear model, and two machine-learning methods---random forests and boosted regression trees. We build our ensembles on a simulated dataset of 5760 fish stocks and test our models with cross-validation and a global database of `r ram_stocks_n` stock assessments. We find that ensemble models substantially improve estimates of population status and trajectory. Machine-learning ensembles performed the best: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, correlation improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the slope of status. When the simulation-trained ensembles were applied to global fish stocks, the best ensembles still improved all dimensions of performance over the next-best individual model. Ensemble models have potential to improve estimates of population status and trajectory across many taxa beyond fish. 
However, ensemble methods must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations they are applied to.

@@

There are often multiple models we can apply to an ecological system to assess population status and trajectory. Models might incorporate different data types, assume alternate population dynamics, or make contrasting assumptions about the starting state of a system. These models may suggest different conclusions about population status and trajectory and we often do not know which model is best. How can we reconcile multiple models to make robust management decisions about ecological resources?

Ensemble methods, which combine the output from multiple models as predictors in a new 'ensemble' model, are one solution to this problem. A simple ensemble model might take the average of multiple model outputs, perhaps weighted by some goodness of fit. A complex ensemble model might also take into account non-linear interactions between model outputs and additional covariates. Ensemble models are popular in other fields including public health [e.g. @thomson2006], agriculture [e.g. @cantelaube2005], climate science [e.g. @murphy2004; @tebaldi2007; @pierce2009], and machine learning [e.g. @dietterich2000]. We think they have tremendous opportunity in ecological resource management.

In fisheries science, a common task is estimating the status and trajectory of an exploited fish population. For the majority of fish stocks, we have limited biological information and stock status isn't known [@fao2014]. In recent years, a number of methods have been proposed to derive population status based on the limited information that is broadly available: fisheries catch time series and basic knowledge about the productivity of the species. Recently, @rosenberg2014 investigated the ability for these models to estimate population status through a large-scale simulation model. A key finding was that these models frequently disagree on population status and any one model has relatively poor predictive ability on average (e.g. Fig. \ref{motivate}).

@@

We first developed and tested ensemble models with a fully factorial simulated dataset [@rosenberg2014]. In summary, the simulation models included three life histories: small pelagic, demersal, large pelagic; three levels of how depleted biomass was at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four exploitation dynamics: a constant exploitation rate, a exploitation rate that is coupled with biomass to mimic an open-access single-species fishery, a scenario where exploitation rate increases to a fixed level, and a 'roller-coaster' scenario where the exploitation rate increases and then decreases. Process noise was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and that process noise was either uncorrelated or had a first-order autoregressive correlation of 0.6. The simulation included a scenario without observation error and with multiplicative observation error around catch at $N(0, 0.2)$.

@@

We also tested our ensemble models on the RAM Legacy Stock Assessment Database [@ricard2012]. Our analysis of the stock-assessment database was based on version XX downloaded on XX. After removing stocks for which the models described below did not converge, this database included `r ram_stocks_n` from XX countries.

@@

We fit four individual data-limited models to estimate \bbmsy\\. Three of the models are mechanistic and based generally on @schaefer1954 biomass dynamics of the form

@@

* *COM-SIR* (catch-only-model with sample-importance-resampling) is a coupled
  harvest-dynamics model developed by @vasconcellos2005. Biomass is assumed to
  follow a Schaefer model and harvest dynamics are assumed to follow a logistic
  model. The model is fit with a sample-importance-sampling algorithm developed
  in @rosenberg2014.

* *CMSY* (catch-MSY) implements a stock-reduction analysis with Schaefer
  biomass dynamics [@martell2013]. It requires a prior distribution for $r$ and
  assigns a prior to the relative proportion of biomass at the end compared to
  unfished biomass (depletion) based on the percentage of maximum catch at the
  end of the time series.

* *SSCOM* (state-space catch-only-model) TODO: Jim -- can you give a ~2 sentence summary?

* *mPRM* (modified panel regression model) is a modified version of the
  panel-regression model used in @costello2012. Unlike the other models
  considered, mPRM is empirical and not mechanistic --- it uses the RAM Legacy
  Stock Assessment database to fit a linear model to a series of
  characteristics of the catch time series and stock with stock-assessed
  \bbmsy\\ as the response. The model used in this paper and @rosenberg2014 is
  modified from the original in that it condenses the life-history categories
  into three categories to match the simulated dataset.
  <!--TODO: what are we doing with the RAM stocks again?-->
  <!--TODO: bias adjustment?-->
  <!--TODO: build in fig:didactic -->

@@

An ensemble model can combine models through anything from a simple average to a complex non-linear model. Here we chose four ensemble models to compare: an average, a linear model with two-way interactions, a random forest, and boosted regression tree. The general form of our models was

@@

\noindent 
where $L$ is a generalized loss function, $f$ is some generalized regression function, $\hat{b}_\mathrm{CMSY}$ is the mean or slope prediction from the CMSY model etc. (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the spectral density of the scaled catch series at frequencies of 0.2 and 0.05, $b$ is the mean or slope value the model is fitted to, and $\theta$ is the set of parameters for the regression function. For all ensemble models of mean \bbmsy\\, which is bounded at zero, we fit the models in log space and exponentiated the predictions to derive a median estimate of \bbmsy\\. For the estimates of \bbmsy\\ slope, we fit ensemble models on the natural untransformed scale.

@@

Applied to the simulated dataset of known status, the individual models had variable success at recovering the mean (status) and slope (trajectory) of \bbmsy\\ in the last five years. All models exhibited a high degree of scatter around the 1:1 line of perfect status prediction (Fig. \ref{hexagon}). CMSY exhibited bimodal predictions of status but had the best rank-order correlation and accuracy scores (Fig. \ref{performance}a). COM-SIR and SSCOM both correctly identified a number of low status stocks, but frequently predicted a high status when status was in fact low (Fig. \ref{hexagon}b, c). mPRM had relatively poor ability to predict status for the simulated dataset (Fig. \ref{hexagon}d). There was generally little correlation between true and predicted trajectory for any of the individual models besides SSCOM (Figs \ref{scatter-sim-slope}a--d, \ref{performance-sim-slope}).

@@

The ensemble models were able to improve the predictions of status by exploiting the best properties of individual models, the covariance between individual models, and interactions with other covariates. For example, SSCOM had strong predictive ability when it predicted low \bbmsy\\ (Fig. \ref{hexagon}c, Fig. \ref{partial-sim}c) and CMSY predictions were approximately linearly related to \bbmsy\\ within the low and high prediction clusters (Fig. \ref{partial-sim}). Ensembles also exploited the covariance between individual model predictions. For instance, both the linear model and GBM ensemble suggest that if mPRM and SSCOM predict high status, the true status also tends to be high (Figs \ref{lm-coefs}, \ref{partial-2d-sim}l). The addition of spectral density covariates helped the ensemble models correctly predict higher status values (Fig. \ref{hexagon}g, h) by exploiting interactions between the spectral density predictors and the individual model predictions (Fig. \ref{partial-2d-sim}). Although the additional covariates improved ensemble fit, performance of the ensembles was only marginally degraded by removing these covariates (Fig. \ref{hexagon} vs. Fig. \ref{hexagon-sim-basic}) TODO NEW FIG THAT QUANTIFIES THIS.
